Tight Generalization Bounds for Hard Parameter Sharing in High-dimensional Linear Regression

### 2020-10-12 arxiv/diff_cov.tex
minor

\paragraph{How does hard parameter sharing scale with sample size and covariate shift?}
One can see that both the variance limit and the bias limit depend heavily on both tasks' samples sizes and the covariate shift matrix $M$.
Next, we illustrate how varying them impact prediction loss, respectively.
\begin{example}[Sample ratio]\label{ex_sample_ratio}
	We first consider the impact of varying sample sizes.
	Consider the random-effects model from Section \ref{sec_same}, with both tasks having an isotropic population covariance matrix.

@@

\begin{corollary}\label{cor_MTL_loss}
	In the setting of Example \ref{ex_sample_ratio}, assume that
	%a) the sample sizes $n_1$ and $n_2$ are greater than $(1 + \tau) p$, b) $\Sigma_1=\Sigma_2=\id_p$, and c) %there exists a small constant $c_0>0$ such that
	(i) both tasks sample sizes are at least $\frac{3}{2}p$;
	(ii) noise variance is smaller than the shared signal variance: $\sigma^2 \lesssim  \kappa^2$;
%	\be\label{choiceofpara0}
%	p^{-1/2+c_0}\sigma^2 + p^{c_0}d^2\le \kappa^2\le p^{1-c_0} (\sigma^2 +d^2)  .  	\ee
	%\be\label{choiceofpara0}
%	(ii) the task-specific variance of $\beta_i$ is much smaller than the signal strength {\color{red}$d^2 = \oo( {\kappa^2})$}; \HZ{what does $\ll$ mean exactly?}
%	(iii) the sample sizes $n_1$ and $n_2$ are greater than $(1 + \tau) p$.
	(iii) task-specific variance is much smaller than the shared signal variance: $d^2 \le p^{-\delta}{\kappa^2}$ for a small constant $\delta>0$.
	Let $\varepsilon = (1 + \sqrt{p/n_1})^ 4 - 1$, which decreases as $n_1$ increases.
	The prediction loss of hard parameter sharing for task two satisfies that
	\begin{align}\label{cor_MTL_error}
	%-\left[1- \left( 1-\frac{1}{\sqrt{\rho_1}}\right)^4\right] pd^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} +\OO(p^{-c}\sigma^2)  \le
	  \left|L(\hat{\beta}_2^{\MTL}) - \frac{d^2 n_1^2 (n_1 + n_2)}{(n_1 + n_2 - p)^3} -\frac{\sigma^2 p}{n_1 + n_2 - p}  \right| \le \varepsilon \cdot \frac{d^2 n_1^2 (n_1 + n_2)}{(n_1 + n_2 - p)^3} + p^{-c } \cdot \OO(d^2 + \sigma^2).
	%\left[\left( 1+\frac{1}{\sqrt{\rho_1}}\right)^4-1\right] d^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} \\
	%& +C \left[(p^{-c_\varphi}+p^{-c_\infty/2})(\sigma^2 +d^2)+p^{-c_\infty}\kappa^2 + %\frac{d^4+\sigma^2 d^2}{\kappa^2}\right],\nonumber
	 \end{align}
	 with high probability for any fixed $c\in(0, \min(\frac{1}{4}, \delta,\frac{\varphi-4}{2\varphi}))$.
%	 {\color{red}[FY: the error also contains $p^{-1/2+2c}\kappa^2 +  p^{-1/4+c} (\sigma^2 +d^2) $, both of which cannot be omitted, because (i) there is no assumption on the upper bound of $\kappa^2$, and (ii) we do not necessarily have $c_\varphi<1/4$. We can decide how to present the result concisely (for instance we can impose an upper bound on $\kappa^2$ and that $c_\varphi<1/4$), but it needs to be correct.]}
	 \end{corollary}

@@

Next, we illustrate an interesting phenomenon where adding task one's samples helps task two initially, but hurts eventually.
Consider the limiting estimate on the left hand side of equation \eqref{cor_MTL_error}.
We vary sample ratio $n_1 / n_2$ by fixing $n_2$ and increasing $n_1$.
%As a function of the sample ratio, the limiting estimate always decreases first from $\frac{\sigma^2 p}{n_2 - p}$ with $n_1$ being zero, and then increases to $d^2$ when $n_1$ goes to infinity.
%We describe a sketch of the proof.
The variance term always reduces as $n_1$ increases.
The bias term always increases as $n_1$ increases, which can be verified by calculating the derivative of the bias term.
\begin{enumerate}
	\item When $\frac{2d^2}{d^2 + \sigma^2} > \frac{p}{n_2 - p}$, the prediction loss decreases monotonically as $n_1$ increases.
	\item When $\frac{2d^2}{d^2 + \sigma^2} < \frac{p}{n_2 - p}$, the limiting estimate always decreases first from $\frac{\sigma^2 p}{n_2 - p}$ with $n_1$ being zero, and then increases to $d^2$ when $n_1$ goes to infinity.
	To see this, near the point where $n_1$ is zero, one can verify that bias increases less than the variance decreases, and there is only one positive root for the derivative being zero.
%Combined together, we observe that the \textit{optimal} sample ratio between $n_1$ and $n_2$ should be among one of following three cases:
%(i) $n_1 = 0$, in which case the prediction loss reduces to $\frac{\sigma^2 p}{n_2 - p}$;
%(ii) $n_1 = \infty$, in which case the prediction loss reduces to $d^2$;
%(iii) the critial point where the derivative of the bias and the variance terms w.r.t. $n_1$ equals zero.
\end{enumerate}

>>>>

\paragraph{How does hard parameter sharing scale with sample size and covariate shift?}
One can see that both the variance limit and the bias limit depend heavily on both tasks' samples sizes and the covariate shift matrix $M$.
Next, we illustrate how varying them impact prediction loss, respectively.
\begin{example}[Sample size ratio]\label{ex_sample_ratio}
	We first consider the impact of varying sample sizes.
	Consider the random-effects model from Section \ref{sec_same}, with both tasks having an isotropic population covariance matrix.

@@

\begin{corollary}\label{cor_MTL_loss}
	In the setting of Example \ref{ex_sample_ratio}, assume that
	%a) the sample sizes $n_1$ and $n_2$ are greater than $(1 + \tau) p$, b) $\Sigma_1=\Sigma_2=\id_p$, and c) %there exists a small constant $c_0>0$ such that
	(i) both tasks sample sizes are at least $3p$;
	(ii) noise variance is smaller than the shared signal variance: $\sigma^2 \lesssim  \kappa^2$;
%	\be\label{choiceofpara0}
%	p^{-1/2+c_0}\sigma^2 + p^{c_0}d^2\le \kappa^2\le p^{1-c_0} (\sigma^2 +d^2)  .  	\ee
	%\be\label{choiceofpara0}
%	(ii) the task-specific variance of $\beta_i$ is much smaller than the signal strength {\color{red}$d^2 = \oo( {\kappa^2})$}; \HZ{what does $\ll$ mean exactly?}
%	(iii) the sample sizes $n_1$ and $n_2$ are greater than $(1 + \tau) p$.
	(iii) task-specific variance is much smaller than the shared signal variance: $d^2 \le p^{-\delta}{\kappa^2}$ for a small constant $\delta>0$.
	Let $\varepsilon = (1 + \sqrt{p/n_1})^ 4 - 1$, which decreases as $n_1$ increases.
	The prediction loss of hard parameter sharing for task two satisfies that
	\begin{align}\label{cor_MTL_error}
	%-\left[1- \left( 1-\frac{1}{\sqrt{\rho_1}}\right)^4\right] pd^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} +\OO(p^{-c}\sigma^2)  \le
	  \left|L(\hat{\beta}_2^{\MTL}) - \frac{d^2 n_1^2 (n_1 + n_2)}{(n_1 + n_2 - p)^3} -\frac{\sigma^2 p}{n_1 + n_2 - p}  \right| \le \varepsilon \cdot \frac{d^2 n_1^2 (n_1 + n_2)}{(n_1 + n_2 - p)^3} + p^{-c } \cdot \OO(d^2 + \sigma^2).
	%\left[\left( 1+\frac{1}{\sqrt{\rho_1}}\right)^4-1\right] d^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} \\
	%& +C \left[(p^{-c_\varphi}+p^{-c_\infty/2})(\sigma^2 +d^2)+p^{-c_\infty}\kappa^2 + %\frac{d^4+\sigma^2 d^2}{\kappa^2}\right],\nonumber
	 \end{align}
	 with high probability for any fixed $c\in(0, \min(\frac{1}{4}, \delta,\frac{\varphi-4}{2\varphi}))$.
%	 {\color{red}[FY: the error also contains $p^{-1/2+2c}\kappa^2 +  p^{-1/4+c} (\sigma^2 +d^2) $, both of which cannot be omitted, because (i) there is no assumption on the upper bound of $\kappa^2$, and (ii) we do not necessarily have $c_\varphi<1/4$. We can decide how to present the result concisely (for instance we can impose an upper bound on $\kappa^2$ and that $c_\varphi<1/4$), but it needs to be correct.]}
	 \end{corollary}

@@

Next, we illustrate an interesting phenomenon where adding task one's samples helps task two initially, but hurts eventually.
Consider the limiting estimate on the left hand side of equation \eqref{cor_MTL_error}.
We vary sample ratio $n_1 / n_2$ by fixing $n_2$ and increasing $n_1$.
%As a function of the sample ratio, the limiting estimate always decreases first from $\frac{\sigma^2 p}{n_2 - p}$ with $n_1$ being zero, and then increases to $d^2$ when $n_1$ goes to infinity.
%We describe a sketch of the proof.
The variance term always reduces as $n_1$ increases.
The bias term always increases as $n_1$ increases, which can be verified by calculating the derivative of the bias term.
\begin{enumerate}
	\item When $\frac{2d^2}{d^2 + \sigma^2} > \frac{p}{n_2 - p}$, \FY{$\frac{d^2}{\sigma^2}<\frac{p}{2n_2-3p} $} the prediction loss decreases monotonically as $n_1$ increases.
	\item When $\frac{2d^2}{d^2 + \sigma^2} < \frac{p}{n_2 - p}$, \FY{$\frac{d^2}{\sigma^2}>\frac{p}{2n_2-3p}$} the limiting estimate always decreases first from $\frac{\sigma^2 p}{n_2 - p}$ with $n_1$ being zero, and then increases to $d^2$ when $n_1$ goes to infinity.
	To see this, near the point where $n_1$ is zero, one can verify that bias increases less than the variance decreases, and there is only one positive root for the derivative being zero.
%Combined together, we observe that the \textit{optimal} sample ratio between $n_1$ and $n_2$ should be among one of following three cases:
%(i) $n_1 = 0$, in which case the prediction loss reduces to $\frac{\sigma^2 p}{n_2 - p}$;
%(ii) $n_1 = \infty$, in which case the prediction loss reduces to $d^2$;
%(iii) the critial point where the derivative of the bias and the variance terms w.r.t. $n_1$ equals zero.
\end{enumerate}
